---
title: "Extracting New York Solar Resource Data from NREL API"
author: "Uni Lee"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(dplyr)
library(tidyverse)
```

# Extracting web data using API in R

## Goal: extract solar resource data for New York City in JSON format and convert it into a dataframe

[Data Gov API](https://developer.nrel.gov/docs/solar/solar-resource-v1/) provides data on solar resource in New York city. We will extract this data as a dataframe in R. 

## Data

The solar resource data provides the following information:

*The Average Direct Normal Irradiance(avg_dni): Direct Normal Irradiance is the amount of solar radiation a surface receives per unit area.
*The Average Tilt at Latitude (avg_lat_tilt): Tilt at Latitude is the amount of radiation a surface receives per unit area that doesn't arrive on a direct path from the sun.
*The Average Global Horizontal Irradiance(avg_ghi): Global Horizontal Irradiance is the total amount of shortwave radiation a surface horizontal to the ground receives from above.

All of the variables above are monthly averages.

## Set parameters for this query

The base url for this API is "https://developer.nrel.gov". Endpoints follow this format: "/api/solar/solar_resource/v1.format?parameters". We will use the json format.

```{r}
# URL
base <- "https://developer.nrel.gov"
endpoint <- "/api/solar/solar_resource/v1.json"
```

To access the database, one must sign up to get a token. I have stored my token in this variable, but I will hide it in the final publication of this document. 

```{r}
# My token
api_key <- "token"

# Other parameters
lat <- 41 # Location info
lon <- -75 

# Save the parameters in a list
params <- list(api_key=api_key, lat=lat, lon=lon)
```

## Query the API

```{r}
url <- modify_url(base, path=endpoint) # specify API url
response <- GET(url, query=params) # query the API using the parameters 

status_code(response) # Check if the query was successful
http_type(response) # Is it the format we want? 

result <- content(response, as="text") # Get content as text
df <- jsonlite::fromJSON(result) # Turn it into a list
```

The result of our query is a complex list. In this list, we subset the the 6th list, which contains information about the variables of our interest.

```{r}
str(df) # Check which list we need to pull
df_4175 <- df[[6]] 

str(df_4175) # Check the structure of the subsetted list
```

Note that even after sub-setting, the result is still nested lists. Since we are only interested in monthly average for each variable (avg_dni, avg_ghi, avg_lat_tilt), we will extract the second list from each list. Then, we can combine them as a whole dataframe. 

```{r}
# Extract the lists as individual vectors
avg_dni <- df_4175$avg_dni$monthly
avg_ghi <- df_4175$avg_ghi$monthly
avg_lat_tilt <- df_4175$avg_lat_tilt$monthly

# Combine the vectors into a dataframe
df_monthly <- tibble::tibble("month" = factor(month.abb, levels=month.abb),
                             "avg_dni" = avg_dni,
                            "avg_ghi" = avg_ghi,
                            "avg_lat_tilt" = avg_lat_tilt)

```

## 3. Convert a complex list into a dataframe

Consider the outputs list, named as an object "df". This is the raw results of our query. Note that this is a complex list of 6, some of which has nested lists of up to 3. We will simplify the outputs list using unlist() and matrix() functions. 

```{r}
str(df)
```
### Workflow: unlist() -> matrix() -> as.data.frame() 

```{r}
# 1. Simplify the list. The unlist() function removes all sub-hierarchies from the list and returns a simple vector.

df_1 <- unlist(df)

# 2. Restructure the vector into a matrix.
df_2 <- matrix(df_1)

# 3. Convert it into a dataframe
df_3 <- as.data.frame(df_2)

```

## 4. Create a function that performs all of the commands above. 

```{r}

endpoint <- "/api/solar/solar_resource/v1.json"

get_nrel <- function(endpoint, params){
  base <- "https://developer.nrel.gov"

  url <- modify_url(base, path=endpoint) # specify API url
  response <- GET(url, query=params) # query the API using the parameters 

  if (http_error(response)) {
    stop("something went wrong", .call=FALSE)
  }
  if (http_type(response) != "application/json") {
    stop("not json!", .call=FALSE)
  }

  result <- content(response, as="text") # Get content as text
  df <- jsonlite::fromJSON(result) # Turn it into a list
  df_1 <- df[[6]]
  
  # Extract the lists as individual vectors
  avg_dni <- unlist(df_1$avg_dni$monthly)
  avg_ghi <- unlist(df_1$avg_ghi$monthly)
  avg_lat_tilt <- unlist(df_1$avg_lat_tilt$monthly)

# Combine the vectors into a dataframe
  df_monthly <- tibble::tibble("month" = factor(month.abb, levels=month.abb),
                             "avg_dni" = avg_dni,
                            "avg_ghi" = avg_ghi,
                            "avg_lat_tilt" = avg_lat_tilt)
  df_monthly
  
}

# Set parameters
params <- list(api_key=api_key, lat=lat, lon=lon)
data <- get_nrel(endpoint, params)
```


## 5. Plot data 

```{r}
data_long <- data %>% pivot_longer(cols=c(avg_dni, avg_ghi, avg_lat_tilt), names_to = "category", values_to="value")

ggplot(data_long, aes(x=month, y=value, group=category, color=category)) + geom_line() + geom_point() + theme_bw() + 
  theme(legend.position="bottom") + 
  labs(x="Month", y="Solar resource (kWh/m^2/day")

```


