---
title: "How skewed is Fandango.com's Movie rating?"
author: "Uni Lee"
date: "2/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(stringr)
```

# Are movie ratings on Fandango.com still inflated? 

The goal of this guided project is to compare Fandango.com's movie ratings in 2015 to those in 2016 to determine if there has been any change. 

In 2015, an investigation by a data jounralist named Walt Hicky from [FiveThirtyEight](https://fivethirtyeight.com/features/fandango-movies-ratings/) revealved that the movie ratings on Fandango is significantly more left-skewed than those of the competitors. After the report was released, Fandango.com announced that it was caused by a bug in the system and that they will fix it shortly. Have they fixed the bug? If they have not, we can assume that the left-skewedness of movie ratings was intentional. 

# Data

To answer the research question, we will compare the frequency distributions of Fandango's movie ratings in 2015 to those in 2016. Walt Hicky has made the data available on [Github](https://github.com/fivethirtyeight/data/tree/master/fandango). The 2016 data has also been made available on [Github](https://github.com/mircealex/Movie_ratings_2016_17) thanks to the Dataquest team. 

```{r}
# Read in data from the repositories
f_2015_raw <- read.csv("data/12_fandango/fandango_score_comparison.csv", sep=";")
f_2016_raw <- read.csv("data/12_fandango/movie_ratings_16_17.csv", sep=";")

# Select relevant variables
f_2015 <- f_2015_raw %>% select(FILM, starts_with("Fandango"))
f_2016 <- f_2016_raw %>% select(movie, year, fandango)

head(f_2015_raw)
head(f_2016_raw)
```

## Sampling errors

To draw general conclusions about the population (all movies that have been rated by Fandango), we would randomly select samples from the population. Since we are comparing the ratings in 2015 to 2016, we need two datasets that consist of randomly selected samples in 2015 and 2016. 

However, the datasets we have are not randomly selected samples of the population. The datasets were selected purposefully according to criteria that the data collectors used for the purpose of their analysis. This puts us at the risk of introducing large sampling errors. 

In order to overcome the limitation of the datasets in our hands, we have to come up with a proxy goal for this analysis. Instead of comparing all movies on Fandango, we will compare ratings of popular movies in 2015 to those of popular movies in 2016. 

# Do the ratings of popular movies in 2015 differ from 2016? 

To compare the two datasets, we have to make sure that the datasets do the following: 

*Each dataset contains only the movies that were released in that year. 
*Movies were selected based on the same criteria for popularity. 

First, we will create two subsets of movies by the release year. 

```{r}
# Although the 2015 dataset does not have a separate column on release year, it is included in the movie title. We can extract this data using stringr package.

only_2015 <- f_2015 %>% mutate(year=str_sub(FILM, -5, -2)) %>% filter(year==2015) # 129 movies

# The 2016 dataset conveniently provides the release year column. 

only_2016 <- f_2016 %>% filter(year==2016) # 191 movies
```

Secondly, we have to make sure that both datasets follow the same criteria for popularity. Unfortunately, the two datasets use different definitions of "popular" movies. Whereas the 2015 dataset contains movies with more than 30 votes, it is unclear how the second dataset selected movies. 

```{r}
# 2015 dataset contains only movies whose votes are over 30.
nrow(only_2015 %>% filter(Fandango_votes<30))

# There is no column for votes in the second dataset
names(only_2016)
```

How do we know that the second dataset has enough popular movies to be representative of 2016? We can randomly sample 10 movies from this dataset and check manually to see if they have received more than 30 votes.

```{r}
set.seed(1)
sample_n(only_2016, size=10)
```

## Change in Fandango's rating system

When I went to Fandango.com to check the votes of the randomly selected movies, I noticed that Fandango no longer uses 5-star rating system. Instead, it uses Audience Scores verified by Rotten Tomatoes. 

* Audience Score: The percentage of users who made a verified movie ticket purchase and rated this 3.5 stars or higher.

To check movies' popularity, we can use the number of reviews on Rotten Tomatoes. However, this would still be imperfect because we don't know how comparable the size of reviewers on Fandango is to that of Rotten Tomatoes. In addition, the number of reviews must have increased since the release date (2016) over time. 

Despite the limitations, we can conclude that the second dataset represents popular movies in 2016.

# Analysis

For our analysis, we will draw two kernel density plots of movie ratings in 2015 and 2016.

```{r}
# Combine two datasets into one table
only_2015_2 <- only_2015 %>% rename("movie" = FILM, "fandango" = Fandango_Stars) %>% select(movie, fandango) %>% mutate(year=2015)
ratings_full <- rbind(only_2015_2, only_2016)
ratings_full$year <- as.factor(ratings_full$year)

# Plot
ggplot(ratings_full, aes(x=fandango, color=year)) + 
  geom_density() +
  labs(title="Distribution of Fandango movie ratings in 2015 - 2016", 
       x="Star Ratings",
       y="Density") +
  xlim(0,5) + 
    theme_classic() +
  theme(plot.title=element_text(hjust=0.5)) +
  scale_color_manual(values=c("red", "blue")) +
  geom_vline(xintercept=mean(only_2015_2$fandango), linetype="dashed", color="red") +
  geom_vline(xintercept=mean(only_2016$fandango), linetype="dashed", color="blue")

```

In this graph, we can observe that the mean rating decreased in 2016. There are less frequencies for ratings above 4.5. However, the ratings are still heavily left-skewed. 

