---
title: "What is the right strategy to win Jeopardy?"
author: "Uni Lee"
date: "4/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(lubridate)
```

Have you ever watched Jeopardy? How about [Wer Wird Millinär](https://en.wikipedia.org/wiki/Wer_wird_Million%C3%A4r%3F_(German_game_show)? They are quiz shows that awards a large sum of money to people who solve trivia questions. To win the game, participants have to stufy a large database of trivia to increase chances of winning. There could be literally infinite number of trivia questions - Who has time to study them all? 

We can find an optimal winning strategy by analyzing historical data. We will identify some patterns in the questions so that we can priortize topics that appear more often. 

Dataquest provided a subset of a large dataset of questions from the US Jeopardy show. The dataset contains 20,000 questions from episodes that aired from September 1984 until January 2012.  


```{r}
jeopardy <- read.csv("data/15_jeopardy.csv") %>% clean_names()
head(jeopardy, 5)
unique(jeopardy$round) # Which round did the question come from?

# Find the time range of episodes
## First, change the column type to an appropriate format
jeopardy <- jeopardy %>% 
  mutate(air_date=ymd(air_date))

max(jeopardy$air_date)
min(jeopardy$air_date)



```




