---
title: "Visualizing Forest Fire Data"
author: "Uni Lee"
date: "12/4/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

Forecasting forest fires is important to minimize environmental and economic damages. [Cortez and Morais](http://www3.dsi.uminho.pt/pcortez/fires.pdf) used a Data Mining approach to predict burned area of forest fires in Portugal. 

In this exercise, we will perform an exploratory analysis of the forest fire dataset via visualization. 

# Data 
## Understanding the raw data structure
```{r}
# Load data from https://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/
fire_raw <- read.csv("data/3_forestfires.csv")
str(fire_raw)
```
Each row represents a forest fire event. There are 13 variables that describe each event. Refer to the following description of each variable, taken from the codebook. 

* X: X-axis spatial coordinate within the Montesinho park map (1 to 9)
* *Y*: Y-axis spatial coordinate within the Montesinho park map (2 to 9)
* *month*: Month of the year: 'jan' to 'dec'
* *day*: Day of the week ('mon' to 'sun')
* *FFMC*: Fine Fuel Moisture Code index from the FWI system (18.7 to 96.20). This code is an indicator of the relative ease of ignition and the flammability of fine fuel.
* *DMC*: Duff Moisture Code index from the FWI system (1.1 to 291.3). It is a numeric rating of the average moisture content of loosely compacted organic layers of moderate depth. This code gives an indication of fuel consumption in moderate duff layers and medium-size woody material.
* *DC*: Drought Code index from the FWI system (7.9 to 860.6). This code is a useful indicator of seasonal drought effects on forest fuels and the amount of smoldering in deep duff layers and large logs.
* *ISI*: Initial Spread Index from the FWI system (0.0 to 56.10). It is a numeric rating of the expected rate of fire spread, calculated based on wind speed and FFMC.
* *temp*: Temperature in Celsius degrees (2.2 to 33.30)
* *RH*: Relative humidity in percentage: 15.0 to 100
* *wind*: Wind speed in km/h: 0.40 to 9.40
* *rain*: Outside rain in mm/m2 : 0.0 to 6.4
* *area*: The burned area of the forest (in ha): 0.00 to 1090.84

*The acronym FWI stands for "fire weather index", a method used by scientists to quantify risk factors for forest fires. Read more [here](https://cwfis.cfs.nrcan.gc.ca/background/summary/fwi)*

## Prepping data for visualization

```{r}
# Convert the month and day into a categorical variable, making sure that the months and day are ordered correctly (rather than in alphabetical order). 

fire <- fire_raw %>% mutate(
  month = factor(month, levels= c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")),
  day = factor(day, levels = c("mon", "tue", "wed", "thu", "fri", "sat", "sun"))
)
```

# Visualization
## Which months and days of the week do forest fires happen the most?

During the months of August and September, forest fires occur the most frequently. 

```{r message=FALSE, warning= FALSE, error=FALSE}
# by month
count_month <- fire %>% group_by(month) %>% 
  summarize(count = n()) %>% as_data_frame()

ggplot(count_month, aes(x=month, y=count)) + 
  geom_col() + 
  labs(title = "Number of forest fired by month") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust=0.5)) 
```

Forest fires are more likely to occur during weekends (fri-sun) than weekdays (mon-thu). 

```{r}
# by day
count_day <- fire %>% group_by(day) %>% 
  summarize(count = n())

ggplot(count_day, aes(x=day, y=count)) + 
  geom_col() + 
  labs(title = "Number of forest fired by day") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust=0.5)) 
```

## Examining time variance of other variables

During the months of Aug-Sept, the following variables exhibit high values, showing positive correlation with frequency of forest fires.
* DC
* DMC
* Temperature

```{r error=FALSE, message=FALSE, warning==FALSE}
# Turn the wide format into a long format
fire_long <- fire %>% pivot_longer(
  cols=c("FFMC", "DMC", "DC", "ISI", "temp", "RH", "wind", "rain", "area"),
  names_to = "var",
  values_to = "value"
)

# Graph it!
## We use boxplot here because there are many outliers.  
ggplot(fire_long, aes(x=month, y=value)) +
  geom_boxplot() +
  facet_wrap(vars(var), scales="free_y") +
  labs(title = "Monthly changes in variables")


```

## Relationship between severity of forest fires and other variables

The area of burned forest is used as a proxy for "severity". The following scatterplots show that DC, FFMC, and temperature are strongly correlated with severity of forest fires. 

```{r}
fire_long_area <- fire %>% pivot_longer(
  cols=c("FFMC", "DMC", "DC", "ISI", "temp", "RH", "wind", "rain"),
  names_to = "var",
  values_to = "value"
)

ggplot(fire_long_area, aes(x=value, y=area)) +
  geom_point() +
  facet_wrap(vars(var), scales="free_x") +
  labs(title = "Relationship between severity (area burned) and other variables")

# There are outliers in "area" variable that are distorting the visualization. To gain a more accurate understanding of the relationship, we will get rid of the outliers and graph again. 

fire_long_area_outlier <- fire_long_area %>% filter(area<300)
ggplot(fire_long_area_outlier, aes(y=area, x=value)) +
  geom_point() +
  facet_wrap(vars(var), scales="free_x") +
  labs(title = "Relationship between severity (area burned) and other variables")

```

# Conclusion

Our exploratory data visualization has revealed that: 

* DC (Drought Code Index) and temperature can be used as a predictor of both frequency and severity of forest fires.

* FFMC (Fine Fuel Moisture Code index) can be used as a predictor of severity of forest fires.

* DMC (Duff Moisture Code index) can be used as a predictor of frequency of forest fires.


